---
title: "Working with the CQ Data"
date: "2019-01-31"
editor_options: 
  chunk_output_type: console
---

# Working with the CQ Data

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(lubridate)
library(socviz)

```

```{r local-functions}

## Calculate age in years, months, or days
calc_age <- function(birth_date, ref_date = Sys.Date(), unit = "year") {

    require(lubridate)

    period <- as.period(interval(birth_date, ref_date),
                        unit = unit)

    switch(unit,
           year = year(period),
           month = month(period),
           day = day(period),
           stop = "Unknown time unit. Choose year, month, or day.")

}

```

## Load it

```{r load}

filenames <- dir(path = "data/clean",
                 pattern = "*.csv",
                 full.names = TRUE)

filenames

data <- filenames %>% map_dfr(read_csv, .id = "congress")

data <- clean_names(data)

```

## Recodes

```{r recoding-1}

## The start and end and other date variables are CQ's and coded 01/03/1945 (i.e. m/d/y)
## We use lubridate's mdy() to convert them to date objects.
data <- data %>% 
  mutate(start = mdy(start), 
         end = mdy(end), 
         born = mdy(born), 
         death = mdy(death))

sessions <- as_tibble(data.frame(congress = 79:116,
                       start_year = seq(1945, 2019, by = 2),
                       end_year = seq(1947, 2021, by = 2)))

sessions <- sessions %>% 
  mutate(start_year = int_to_year(start_year, "01", "03"),
         end_year = int_to_year(end_year, "01", "03"))


data <- data %>% 
  mutate(congress = as.numeric(congress) + 78)

data
```


```{r jwervin}

data$born[data$last == "Ervin" & data$first == "Joseph"] <- ymd("1901-03-01")

```

Join sessions to main data.

```{r sessions}
data <- left_join(data, sessions)


## Character Vector of all name-dob combinations 
data <- data %>% 
  mutate(name_dob = paste(last, middle, first, born, sep = "--"))

## Uniques
name_dob <- as.character(unique(data$name_dob))

## create pid for unique name_dobs
persons <- tibble(pid = 1:length(name_dob),
                  name_dob = name_dob)

data <- left_join(data, persons)

data <- data %>% 
  mutate(start_age = calc_age(born, start_year))

data %>% 
  select(start_age)

```


William Joseph Green Jr and William Joseph Green III are father
and son; the latter succeeded to his father's seat. The raw data
have a mistaken entry for the son in the 79th (1945) congress,
giving his `start_age` as 6.

```{r wjgreen}

ind <- which(with(data, congress == 79 & ## 1945
                        last == "Green" &
                        middle == "Joseph" &
                        first == "William" &
                        suffix == "III")) ## but this is the son

data <- data[-ind,]


```

## New Variables

- `poc`: Person of color; binary
- `days_old`: N days old
- `months_old`: N months old
- `full_name`: Full Name
- `end_career`: Date of end career
- `entry_age`:  Age when first entered congress
- `yr_fac`: start_year as a factor rather than a date

```{r newvariables}

data <- data %>% 
  mutate(poc = recode(race, "White" = "White", .default = "Non-White"),
         days_old = calc_age(born, start_year, "day"),
         months_old = calc_age(born, start_year, "month"),
         full_name = paste(first, last, suffix),
         full_name = stringr::str_remove(full_name, " NA$"),
         entry_age = calc_age(born, start),
         yr_fac = factor(year(start_year)))


```

```{r}

mean_age <- data %>% filter(position == "U.S. Representative") %>%
    group_by(congress, sex) %>%
    summarize(N = n(), 
              mean_age = mean(start_age, na.rm = TRUE), 
              SD = sd(start_age),
              lo = mean_age - SD, 
              hi = mean_age + SD)

data %>% filter(position == "U.S. Representative",
                    party %in% c("Democrat", "Republican")) %>%
    group_by(congress, sex, party) %>%
    summarize(year = first(start_year), 
              mean_age = mean(start_age, na.rm = TRUE)) %>%
    filter(congress == 100)


```

```{r}

mean_age <- data %>% filter(position == "U.S. Representative") %>%
    group_by(congress) %>%
    summarize(year = first(start_year), 
              mean_age = mean(start_age, na.rm = TRUE))

mean_age
```


```{r}

data %>% filter(position == "U.S. Representative",
                    party %in% c("Democrat", "Republican")) %>%
    group_by(congress) %>%
    summarize(year = first(start_year), 
              mean_age = mean(start_age, na.rm = TRUE)) %>%
    filter(congress == 116)

```

```{r}

data %>% filter(position == "U.S. Representative",
                    party %in% c("Democrat", "Republican")) %>%
    group_by(congress, party) %>%
    summarize(year = first(start_year), 
              mean_age = mean(start_age, na.rm = TRUE)) %>%
    filter(congress == 116)

```

```{r}

median_age <- data %>% 
  filter(position == "U.S. Representative") %>%
  group_by(congress) %>%
    summarize(year = first(start_year), 
              median_age = median(start_age))

median_age_party <- data %>% 
  filter(position == "U.S. Representative") %>%
  group_by(congress, party) %>%
    summarize(year = first(start_year), 
              median_age = median(start_age)) %>%
    filter(party %in% c("Democrat", "Republican"))

median_age_party

```


```{r}
oldest_group_by_year <- data %>% 
  filter(party %in% c("Democrat", "Republican"),
         position == "U.S. Representative") %>%
  group_by(congress, party) %>% 
  filter(start_age > quantile(start_age, 0.99, na.rm = TRUE))

youngest_group_by_year <- data %>% 
  filter(party %in% c("Democrat", "Republican"),
         position == "U.S. Representative") %>%
  group_by(congress, party) %>%
  filter(start_age < quantile(start_age, 0.01, na.rm = TRUE))

```

```{r}

name_ranks <- data %>%
    select(start_year, first, sex) %>%
    filter(sex == "M", .preserve = FALSE) %>%
    group_by(start_year, first) %>%
    summarize(n = n()) %>%
    mutate(pct = (n/sum(n))*100) %>%
    group_by(start_year) %>%
    top_n(3, wt = pct)

name_ranks

```


```{r}
job_ranks <- data %>%
    select(start_year, job_type1) %>%
    group_by(start_year, job_type1) %>%
    summarize(n = n()) %>%
    mutate(pct = (n/sum(n))*100) %>%
    group_by(start_year) %>%
    top_n(3, wt = pct) %>% 
    arrange(desc(pct), .by_group = TRUE)  

job_ranks

```

```{r}

senate <- data %>% 
  filter(position == "U.S. Senator") %>% 
  group_by(pid) %>% 
  summarize(first = first(first), 
            last = first(last), 
            party = first(party),
            state = first(state), 
            start = first(start), 
            end = first(end))

house <- data %>% 
  filter(position == "U.S. Representative") %>% 
  group_by(pid) %>% 
  summarize(state = first(state),
            district = first(district),
            start = first(start), 
            end = first(end))


sen_and_house <- inner_join(senate, house, by = "pid")

sen_and_house

```

