---
title: "More on dplyr"
date: "2019-02-07"
editor_options: 
  chunk_output_type: console
---

# More on dplyr

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(lubridate)
library(socviz)
library(gapminder)

```

## Scoping

```{r scope1}

organdata %>%
  group_by(world) %>%
  summarize_if(is.numeric, mean, na.rm = TRUE) %>%
  select(world, donors, pubhealth, roads) %>%
  select_all(tools::toTitleCase)

```

```{r scope2}

organdata %>%
  group_by(country) %>%
  summarize_if(is.numeric, 
               funs(avg = mean, sd = sd), 
               na.rm = TRUE) %>%
  select(country, donors_avg, 
         donors_sd, roads_avg, roads_sd) %>%
  arrange(desc(donors_avg))

```

```{r scope3}

out <- lm(donors ~ pop + gdp + roads, data = organdata)

summary(out)

```

```{r mapping}

names(summary(out))

```

```{r mapping1}
organdata %>%
  split(.$world) %>%
  map(~ lm(donors ~ pop + gdp + roads, data = .)) %>%
  map(summary) %>%
  map_dbl("r.squared")
```

## Zero Counts in Dplyr

Grab the data remotely, to make this self-contained.

```{r data}

url <- "https://datamgmt.org/data/first_terms.csv"

df <- read_csv(url)

df

```

```{r }
df %>%
    group_by(start_year, party, sex) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N))
```


```{r }
## Hex colors for sex
sex_colors <- c("#E69F00", "#993300")

## Hex color codes for Dem Blue and Rep Red
party_colors <- c("#2E74C0", "#CB454A")

## Group labels
mf_labs <- tibble(M = "Men", F = "Women")

theme_set(theme_minimal())
```

```{r }
df %>%
    group_by(start_year, party, sex) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N)) %>%
    ggplot(aes(x = start_year,
               y = freq,
               fill = sex)) +
    geom_col() +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = sex_colors, labels = c("Women", "Men")) +
    labs(x = "Year", y = "Percent", fill = "Group") +
    facet_wrap(~ party)

```

```{r }
df %>%
    group_by(start_year, party, sex) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N)) %>%
    ggplot(aes(x = start_year,
               y = freq,
               color = sex)) +
    geom_line(size = 1.1) +
    scale_y_continuous(labels = scales::percent) +
    scale_color_manual(values = sex_colors, labels = c("Women", "Men")) +
    guides(color = guide_legend(reverse = TRUE)) +
    labs(x = "Year", y = "Percent", color = "Group") +
    facet_wrap(~ party)
```

### Option 1: recode to factors

```{r }
df_f <- df %>% modify_if(is.character, as.factor)

df_f %>%
    group_by(start_year, party, sex) %>%
    tally()


```

## Option 2: `ungroup()` and `complete()`

```{r }
df %>%
    group_by(start_year, party, sex) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N)) %>%
    ungroup() %>%
    complete(start_year, party, sex,
             fill = list(N = 0, freq = 0))

```

```{r }
df_f %>%
    group_by(start_year, party, sex) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N)) %>%
    ggplot(aes(x = start_year,
               y = freq,
               color = sex)) +
    geom_line(size = 1.1) +
    scale_y_continuous(labels = scales::percent) +
    scale_color_manual(values = sex_colors, labels = c("Women", "Men")) +
    guides(color = guide_legend(reverse = TRUE)) +
    labs(x = "Year", y = "Percent", color = "Group") +
    facet_wrap(~ party)
```

## Spreading multiple values

```{r }
edu
```

```{r }
edu_t <- gather(data = edu,
                key = school,
                value = freq,
                elem4:coll4)

head(edu_t) 


tail(edu_t)
```

### Generate some sample stratified data

```{r }
gen_cats <- function(x, N = 1000) {
    sample(x, N, replace = TRUE)
}

set.seed(101)
N <- 1000

income <- rnorm(N, 100, 50)

vars <- list(stratum = c(1:8),
          sex = c("M", "F"),
          race =  c("B", "W"),
          educ = c("HS", "BA"))

df <- as_tibble(map_dfc(vars, gen_cats))
df <- add_column(df, income)

df
```

### Basic grouped summary (no gathering)

```{r }
tv_wide1 <- df %>% group_by(sex, race, stratum, educ) %>%
    summarize(mean_inc = mean(income), N = n())

tv_wide1

```

### Gather mean income and N into a key-value pair

```{r }
tv_wide2 <- df %>% group_by(sex, race, stratum, educ) %>%
    summarize(mean_inc = mean(income), N = n()) %>%
    gather(variable, value, -(sex:educ))

tv_wide2
```

### Unite the `variable` (key) column with the Education measure

```{r }
tv_wide2 <- df %>% group_by(sex, race, stratum, educ) %>%
    summarize(mean_inc = mean(income), N = n()) %>%
    gather(variable, value, -(sex:educ)) %>%
    unite(temp, educ, variable)

tv_wide2

```

### Finally, spread the resulting temporary column, to get the desired result

```{r }
tv_wide2 <- df %>% group_by(sex, race, stratum, educ) %>%
    summarize(mean_inc = mean(income), N = n()) %>%
    gather(variable, value, -(sex:educ)) %>%
    unite(temp, educ, variable) %>%
    spread(temp, value)

tv_wide2

```

### A function to do this gather, spread, unite sequence

```{r }
multi_spread <- function(df, key, value) {
    # quote key
    keyq <- rlang::enquo(key)
    # break value vector into quotes
    valueq <- rlang::enquo(value)
    s <- rlang::quos(!!valueq)
    df %>% gather(variable, value, !!!s) %>%
        unite(temp, !!keyq, variable) %>%
        spread(temp, value)
}

```


### Use the function directly

```{r }
## Final version
tv_wide3 <- df %>% group_by(sex, race, stratum, educ) %>%
    summarize(mean_inc = mean(income), N = n()) %>%
    multi_spread(educ, c(mean_inc, N))

tv_wide3

```

## Nesting

```{r }
out_le <- gapminder %>%
    group_by(continent, year) %>%
    nest()

out_le

```

```{r }
out_le %>% filter(continent == "Europe" & year == 1977) %>% 
           unnest()

```

```{r }
fit_ols <- function(df) {
    lm(lifeExp ~ log(gdpPercap), data = df)
}

out_le <- gapminder %>%
    group_by(continent, year) %>%
    nest() %>% 
    mutate(model = map(data, fit_ols)) 


out_le
```

```{r }
library(broom) ## more later

fit_ols <- function(df) {
    lm(lifeExp ~ log(gdpPercap), data = df)
}

out_tidy <- gapminder %>%
    group_by(continent, year) %>%
    nest() %>% 
    mutate(model = map(data, fit_ols),
           tidied = map(model, tidy)) %>%
    unnest(tidied, .drop = TRUE)

out_tidy
```

```{r }

out_tidy <- gapminder %>%
    group_by(continent, year) %>%
    nest() %>% 
    mutate(model = map(data, fit_ols),
           tidied = map(model, tidy)) %>%
    unnest(tidied, .drop = TRUE) %>% 
    filter(term %nin% "(Intercept)")

out_tidy
           

```


## Misc dplyr

### Selection

```{r }


organdata %>% 
  select(matches("_lag"))

```

```{r }
organdata %>% 
  select(world, everything())
```

### Summaries

```{r }
organdata %>% 
  summarize_all(funs(min, max), na.rm = TRUE)
```

```{r }
organdata %>% 
  summarize_if(is.numeric, funs(min, max), na.rm = TRUE)
```


```{r }
organdata %>%
  group_by(country) %>%
  summarize_if(is.numeric, funs(min, max), na.rm = TRUE)
```

### Scoped filtering

```{r }
organdata %>% 
  filter_all(any_vars(is.na(.)))
```

### Rowwise operations

```{r }
iris %>% 
     select(contains("Length")) %>%
     rowwise() %>%
     mutate(avg_length = 
              mean(c(Petal.Length, Sepal.Length)))

```
